WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2018-12-13 paper/ms.tex
update text w/ final(?) values

% TODO: final numbers
To verify that the measured uncertainties are plausible, we computed
the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
transit times.  We found that $\chi^2 = 8.4$, with 16 degrees of
freedom.  Visually inspecting residuals showed that the error variance
had been overestimated, so we multiplied the measured TESS errors by a
factor $f=0.73$, forcing a reduced $\chi^2$ of unity.  This lowered
the mean midtime uncertainty for each transit from 30 to 22 seconds.

% TODO: final numbers
Figure~\ref{fig:lightcurves} shows the lightcurves, best-fit models,
and residuals.  Table~2 reports the mid-transit times and their
uncertainties.
Binning the residuals to 1-hour windows and taking the median absolute
deviation, the lightcurves have an RMS of $586\,{\rm ppm}\,{\rm
hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  At
$T\approx11.8$, with a 9-pixel aperture read noise is expected to
contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
in planet detection simulations.  This zodiacal noise prediction seems
to have been an overestimate, given the observed RMS of WASP-4.  A
broader assessment of TESS's photometric performance is beyond
the scope of this work.

@@

%TODO: final numbers
Figure~\ref{fig:times} shows the residuals with respect to the linear
model.  The best linear fit has $\chi^2 = 149$ and 58 degrees of
freedom.  The best quadratic fit has $\chi^2 = 48.0$ and 57
degrees of freedom, compared to $\chi^2 = 51.0$
and 55 degrees of freedom for the best-fit precession model.

%TODO: final numbers. note epoch changes...
The difference in $\chi^2$ between the linear and quadratic fit
corresponds to $p \approx 10^{-23}$.  A more likely possibility is
clearly that a systematic error is affecting one of the most important
sets of times: either the earliest times, the Huitson et al.\ Gemini
times (epoch numbers 44, 322, 591, and 851), or the TESS times.  We
have taken every precaution against systematic offsets in the TESS
timestamps (Appendix~\ref{sec:verify_tess}). Huitson et al.\ stand by
their reported times (J.-M. D\'esert, priv.\
comm.)\footnote{\citet{huitson_gemini_2017} corrected the times to be
at mid-exposure. They also analyzed other hot Jupiters, which all
showed mid-transit times as expected from a constant ephemeris.}.
Since the linear model provides a poor fit to the data, we discard it
from further consideration.

%TODO: final numbers
The quadratic model provides a better fit to the data than the
precession model.  It is favored by $\Delta \chi^2 = 3.0$, and has two
fewer free parameters.  Another useful heuristic for model comparison
is the Bayesian Information Criterion (BIC),
\begin{equation}
  {\rm BIC} = \chi^2 + k\log n,
\end{equation}
for $k$ the number of free parameters, and $n$ the number of data
points. For us, $n=60$.  The difference in the BIC for the precession
and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm
BIC}_{\rm quad} = 11$, corrsonding to a Bayes factor $\approx
4\times10^{4}$.  The Akaike Information Criterion similarly favors the
orbital decay model, by $\Delta {\rm AIC} = 7$.  This is strong
evidence that the quadratic model describes the data better than the
precession model \citep{kass_bayes_1995}.

%TODO: final numbers
If we assume the orbital decay model, we find a period derivative 
\begin{equation}
\dot{P}
  = - (3.82 \pm 0.38)\times 10^{-10}
  = - (12.1) \pm 1.2\,{\rm ms}\,{\rm yr}^{-1}.
  \label{eq:dP_dt_obs}
\end{equation}
For comparison, the period derivative seen by
\citet{maciejewski_departure_2016} and \citet{patra_2017} in WASP-12b
is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  If both planets
are decaying, WASP-4b would be inspiraling at about half the rate as
WASP-12b.

%TODO: final numbers
If we instead assume the precession model, the best-fit eccentricity
is
\begin{equation}
  e = (1.63^{+ 1.48}_{- 0.65})\times10^{-3}
\end{equation}
The longitude of periapse correspondingly advances by $\dot{\omega}
= 14.8^{+5.4}_{-3.8}\ {\rm degrees}\,{\rm yr}^{-1}$, and the
precession period would be $24^{+8}_{-6}$ years.

@@

For the sake of argument, let us assume that the timing variation is
caused entirely by orbital decay.  If the orbital period decreases
at a fixed rate, then it will reach zero after
\begin{equation}
  \frac{P}{ \dot{P} } = 9.6 \, {\rm Myr}.
\end{equation}
For comparison, the survival time found by \citet{patra_2017} for
WASP-12b was 3.2\,Myr.

@@

If the planet's spin and orbit are synchronized and the star is slowly
rotating, then the semi-major axis and eccentricity evolve as
\citep[Appendix B of][]{metzger_optical_2012}
\begin{align}
  \frac{1}{\tau_{\rm e}} &=
  \frac{|\dot{e}|}{e} =
    \frac{63 \pi } {2 Q_{\rm p}' }
    \left( \frac{R_{\rm p}}{a} \right)^5
    \left( \frac{M_\star}{M_{\rm p}} \right)
    \left( \frac{1}{P} \right)
  \label{eq:de_dt}
  \\
  \frac{1}{\tau_{\rm a}} &=
  \frac{|\dot{a}|}{a} =
    \frac{9 \pi } {Q_\star' }
    \left( \frac{R_\star}{a} \right)^5
    \left( \frac{M_{\rm p}}{M_\star} \right)
    \left( \frac{1}{P} \right).
  \label{eq:da_dt}
\end{align}
The orbital period evolves as
\begin{equation}
\label{eq:dP_dt}
  \dot{P} = -\frac{27\pi}{2 Q_\star'}
            \left(\frac{M_{\rm p}}{M_\star}\right)
            \left(\frac{R_\star}{a}\right)^5.
\end{equation}
The modified quality factor of WASP-4 corresponding to the observed
value of $\dot{P}$ is then
\begin{equation}
	Q_\star' \approx 3.3\times10^4,
\end{equation}
which is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
10^5$, based on the orbits of the Galilean moons
\citep{lainey_strong_2009}.  For stars, studies of the binary
eccentricity distribution have found $Q_\star' \approx 10^5 - 10^7$
\citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
geller_direct_2013,milliman_wiyn_2014}.  The dissipation rates may
differ in hot Jupiter systems, since the tidal forcing frequencies and
amplitudes are different.  Observationally-driven studies of hot
Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, from quite
different models
\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,penev_empirical_2018,cameron_hierarchical_2018}.
For instance, motivated by the rapid rotation of some hot Jupiter
hosts \citep{pont_empirical_2009,penev_hats-18b_2016},
\citet{penev_empirical_2018} modeled the evolution of hot Jupiter
systems under the influence of a magnetized wind and a constant
phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
(1.2^{+1.0}_{-0.5})\times10^7$.  This strongly disagrees with the
quality factor we infer from the apparent period change.

@@

Let us instead assume that the timing variation is a portion of an
apsidal precession cycle.  The required orbital eccentricity is then
$e\approx 1.6\times10^{-3}$, and one full precession period would take
about 24 years.

@@

\citet{ragozzine_probing_2009} also noted that if the planet has
non-zero eccentricity, a measurement of $\dot{\omega}$ can be used to
measure the planet's Love number.  From their Equation 14, the implied
Love number for WASP-4b is
\begin{equation}
k_{2,{\rm p}} = 1.6^{+2.1}_{-1.2}.
\end{equation}
For comparison, $k_{2,{\rm Jup}} \approx 0.55$
\citep{wahl_tidal_2016,ni_empirical_2018}.  A uniform density sphere
has $k_2 = 1.5$.  Our uncertainties on $k_{2,{\rm p}}$ are large
because the eccentricity, reference time, and ${\rm d}\omega/{\rm d}E$
are correlated, and the measured occultation times only barely
constrain them (Figure~\ref{fig:future}).

@@

Could the eccentricity truly be non-zero?  For WASP-4b, the
eccentricity damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e =
0.29 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system is well over
$1\,{\rm Gyr}$ old \citep[{\it e.g.},][]{winn_transit_2009}, so any
initial eccentricity should have been erased.

@@

A different way to excite the hot Jupiter's eccentricity is the
Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
In this case, the orbital plane of the outer companion, ``c'', would
need to be inclined relative to that of the hot Jupiter, ``b'', by at
least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
%\citet{bailey_understanding_2019} point out how the joint requirements
%of Kozai-Lidov oscillations with a non-detection of an additional RV
%signal can be combined to produce lower and upper limits on the
%companion's mass.
For the Kozai-Lidov mechanism to operate at maximum efficiency, we
need \citep[][Equation 20]{bailey_understanding_2019}
\begin{equation}
  M_{\rm c} > 7.7\,M_\oplus
  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
  \label{eq:kozai_bound}
\end{equation}
where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
of WASP-4b and the hypothetical WASP-4c, and we have assumed WASP-4b's
Love number is $k_{2,{\rm b}}\approx 0.6$.  For the RV signal of the
companion to remain undetected, it would need to be in the residual
$({\rm O-C})_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
\citet{triaud_spin-orbit_2010}.  Again following
\citet{bailey_understanding_2019}, this implies
\begin{align}
  M_{\rm c} &<
  ({\rm O-C})_{\rm RV}
  \left( \frac{M_\star a_b}{G} \right)^{1/2}
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2}
  \nonumber
  \\
  M_{\rm c} &< 
  24.2\,M_\oplus
  \times 
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2},
  \label{eq:rv_bound}
\end{align}
for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto \sin^2 i_{\rm c}$
a geometric prefactor that depends on the argument of periastron
$\omega_{\rm c}$ and inclination $i_{\rm c}$ of the exterior companion
(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b
transits along the line of sight, $f$ can be arbitrarily small, and
both of the preceding limits can be satisfied.

@@

Another mechanism to produce the observed period change would be if an
unseen companion in the WASP-4 system accelerated the star towards our
line of sight.
This would produce a period derivative
\begin{equation}
	\dot{P} \approx \frac{\dot{v}_{\rm r} P}{c},
\end{equation}
for $\dot{v}_{\rm r}$ the time derivative of the radial velocity.
%Recently, \citet{bonomo_gaps_2017} collected all available HARPS data
%for the system, and reported no detectable slope at $3\sigma$.
%However,
\citet{knutson_friends_2014} collected 5 Keck HIRES radial
velocity measurements for WASP-4, and merged them with measurements
from CORALIE and HARPS \citep{wilson_wasp-4b_2008,pont_determining_2011,husnoo_observational_2012}.
Using these datasets, \citet{knutson_friends_2014} found 
\begin{equation}
\dot{v}_{\rm r} =
   -0.0099^{+0.0052}_{-0.0054}
   \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1},
\end{equation}
or that the
star can be accelerating towards us at most by $\dot{v}_{\rm r} > -
0.021 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ (2$\sigma$).
The associated $\dot{P}$ must be less than $9\times 10^{-11}$.  
Radial acceleration could thus cause at most one-quarter of the
observed period decrease (Equation~\ref{eq:dP_dt_obs}).
Improved radial velocity
monitoring will help clarify the situation,
particularly given that WASP-4 shows short-term RV trends with
%$\dot{\gamma} \approx 2.8 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ over
amplitudes $\sim$$\mathcal{O}(10\,{\rm m\,s}^{-1})$, over tens of days
\citep{husnoo_observational_2012}.  Given that WASP-4 is a
magnetically active star \citep{sanchis-ojeda_starspots_2011}, these
short-timescale variations might conspire to produce spurious
RV signals over long baselines.

>>>>

To verify that the measured uncertainties are plausible, we computed
the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
transit times.  We found that $\chi^2 = 9.2$, with 16 degrees of
freedom.  Visually inspecting residuals showed that the error variance
had been overestimated, so we multiplied the measured TESS errors by a
factor $f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered
the mean uncertainty of the transit midtimes from $29.8$ to $22.6$
seconds.

Figure~\ref{fig:lightcurves} shows the lightcurves, best-fit models,
and residuals.  Table~2 reports the mid-transit times and their
uncertainties.
Binning the residuals to 1-hour windows and taking the median absolute
deviation, the lightcurves have an RMS of $586\,{\rm ppm}\,{\rm
hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  At
$T\approx11.8$, with a 9-pixel aperture read noise is expected to
contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
in planet detection simulations.  This zodiacal noise prediction seems
to have been an overestimate, given the observed RMS of WASP-4.  A
broader assessment of TESS's photometric performance is beyond
the scope of this work.

@@

%TODO: final numbers
Figure~\ref{fig:times} shows the residuals with respect to the linear
model.  The best linear fit has $\chi^2 = 174$ and 61 degrees of
freedom.  The best quadratic fit has $\chi^2 = 62.6$ and 60
degrees of freedom, compared to $\chi^2 = 64.4$
and 58 degrees of freedom for the best-fit precession model.

%TODO: final numbers. note epoch changes...
The difference in $\chi^2$ between the linear and quadratic fit
corresponds to $p \approx 10^{-26}$.  A more likely possibility is
clearly that a systematic error is affecting one of the most important
sets of times: either the earliest times, the Huitson et al.\ Gemini
times (epoch numbers 30, 308, 577, and 837), or the TESS times.  We
have taken every precaution against systematic offsets in the TESS
times (Appendix~\ref{sec:verify_tess}), and Huitson et al.\ stand by
their reported times (J.-M. D\'esert, priv.\
comm.)\footnote{\citet{huitson_gemini_2017} corrected the times to be
at mid-exposure. They also analyzed other hot Jupiters, which all
showed mid-transit times as expected from a constant ephemeris.}.
Since the linear model provides a poor fit to the data, we discard it
from further consideration.

The quadratic model provides a slightly better fit to the data than the
precession model.  It is favored by $\Delta \chi^2 = 1.8$, and has two
fewer free parameters.  Another useful heuristic for model comparison
is the Bayesian Information Criterion (BIC),
\begin{equation}
  {\rm BIC} = \chi^2 + k\log n,
\end{equation}
for $k$ the number of free parameters, and $n$ the number of data
points. For us, $n=62$.  The difference in the BIC for the precession
and decay models is $\Delta {\rm BIC} = {\rm BIC}_{\rm prec} - {\rm
BIC}_{\rm quad} = 10$, corrsonding to a Bayes factor $\approx
1.2\times10^{4}$.  The Akaike Information Criterion similarly favors
the orbital decay model, by $\Delta {\rm AIC} = 5.8$.  This is strong
evidence that the quadratic model describes the data better than the
precession model \citep{kass_bayes_1995}.

If we assume the orbital decay model, we find a period derivative 
\begin{equation}
\dot{P}
  = - (3.98 \pm 0.38)\times 10^{-10}
  = - (12.6) \pm 1.2\,{\rm ms}\,{\rm yr}^{-1}.
  \label{eq:dP_dt_obs}
\end{equation}
For comparison, the period derivative seen by
\citet{maciejewski_departure_2016} and \citet{patra_2017} in WASP-12b
is $\dot{P} = -29 \pm 3\,{\rm ms}\,{\rm yr}^{-1}$.  If both planets
are decaying, WASP-4b would be inspiraling at about half the rate as
WASP-12b.

If we instead assume the precession model, the best-fit eccentricity
is
\begin{equation}
  e = (1.83^{+ 1.88}_{- 0.75})\times10^{-3}
\end{equation}
The longitude of periapse correspondingly advances by $\dot{\omega} =
14.0^{+4.9}_{-3.5}\ {\rm degrees}\,{\rm yr}^{-1}$, and the precession
period would be $26^{+9}_{-7}$ years.

@@

For the sake of argument, let us assume that the timing variation is
caused entirely by orbital decay.  If the orbital period decreases
at a fixed rate, then it will reach zero after
\begin{equation}
  \frac{P}{ \dot{P} } = 9.2 \, {\rm Myr}.
\end{equation}
For comparison, the survival time found by \citet{patra_2017} for
WASP-12b was 3.2\,Myr.

@@

If the planet's spin and orbit are synchronized and the star is slowly
rotating, then the semi-major axis and eccentricity evolve as
\citep[Appendix B of][]{metzger_optical_2012}
\begin{align}
  \frac{1}{\tau_{\rm e}} &=
  \frac{|\dot{e}|}{e} =
    \frac{63 \pi } {2 Q_{\rm p}' }
    \left( \frac{R_{\rm p}}{a} \right)^5
    \left( \frac{M_\star}{M_{\rm p}} \right)
    \left( \frac{1}{P} \right)
  \label{eq:de_dt}
  \\
  \frac{1}{\tau_{\rm a}} &=
  \frac{|\dot{a}|}{a} =
    \frac{9 \pi } {Q_\star' }
    \left( \frac{R_\star}{a} \right)^5
    \left( \frac{M_{\rm p}}{M_\star} \right)
    \left( \frac{1}{P} \right).
  \label{eq:da_dt}
\end{align}
The orbital period evolves as
\begin{equation}
\label{eq:dP_dt}
  \dot{P} = -\frac{27\pi}{2 Q_\star'}
            \left(\frac{M_{\rm p}}{M_\star}\right)
            \left(\frac{R_\star}{a}\right)^5.
\end{equation}
The modified quality factor of WASP-4 corresponding to the observed
value of $\dot{P}$ is then
\begin{equation}
	Q_\star' \approx 2.9\times10^4,
\end{equation}
which is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
10^5$, based on the orbits of the Galilean moons
\citep{lainey_strong_2009}.  For stars, studies of the binary
eccentricity distribution have found $Q_\star' \approx 10^5 - 10^7$
\citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
geller_direct_2013,milliman_wiyn_2014}.  The dissipation rates may
differ in hot Jupiter systems, since the tidal forcing frequencies and
amplitudes are different.  Observationally-driven studies of hot
Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, from quite
different models
\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,penev_empirical_2018,cameron_hierarchical_2018}.
For instance, motivated by the rapid rotation of some hot Jupiter
hosts \citep{pont_empirical_2009,penev_hats-18b_2016},
\citet{penev_empirical_2018} modeled the evolution of hot Jupiter
systems under the influence of a magnetized wind and a constant
phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
(1.2^{+1.0}_{-0.5})\times10^7$.  This strongly disagrees with the
quality factor we infer from the apparent period change.

@@

Let us instead assume that the timing variation is a portion of an
apsidal precession cycle.  The required orbital eccentricity is then
$e\approx 1.8\times10^{-3}$, and one full precession period would take
about 24 years.

@@

\citet{ragozzine_probing_2009} also noted that if the planet has
non-zero eccentricity, a measurement of $\dot{\omega}$ can be used to
measure the planet's Love number.  From their Equation 14, the implied
Love number for WASP-4b is
\begin{equation}
k_{2,{\rm p}} = 1.5^{+2.1}_{-1.1}.
\end{equation}
For comparison, $k_{2,{\rm Jup}} \approx 0.55$
\citep{wahl_tidal_2016,ni_empirical_2018}.  A uniform density sphere
has $k_2 = 1.5$.  Our uncertainties on $k_{2,{\rm p}}$ are large
because the eccentricity, reference time, and ${\rm d}\omega/{\rm d}E$
are correlated, and the measured occultation times only barely
constrain them (Figure~\ref{fig:future}).

@@

Could the eccentricity truly be non-zero?  For WASP-4b, the
eccentricity damping timescale (Equation~\ref{eq:de_dt}) is $\tau_e =
0.30 (Q_{\rm p}'/10^5) \, {\rm Myr}$.  The system is well over
$1\,{\rm Gyr}$ old, so unless the planet arrived very recently, any
initial eccentricity should have been erased.

@@

A different way to excite the hot Jupiter's eccentricity is the
Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
In this case, the orbital plane of the outer companion, ``c'', would
need to be inclined relative to that of the hot Jupiter, ``b'', by at
least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
%\citet{bailey_understanding_2019} point out how the joint requirements
%of Kozai-Lidov oscillations with a non-detection of an additional RV
%signal can be combined to produce lower and upper limits on the
%companion's mass.
For the Kozai-Lidov mechanism to operate at maximum efficiency, we
need \citep[][Equation 20]{bailey_understanding_2019}
\begin{equation}
  M_{\rm c} > 7.1\,M_\oplus
  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
  \label{eq:kozai_bound}
\end{equation}
where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
of WASP-4b and the hypothetical WASP-4c, and we have assumed WASP-4b's
Love number is $k_{2,{\rm b}}\approx 0.6$.  For the RV signal of the
companion to remain undetected, it would need to be in the residual
$({\rm O-C})_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
\citet{triaud_spin-orbit_2010}.  Again following
\citet{bailey_understanding_2019}, this implies
\begin{align}
  M_{\rm c} &<
  ({\rm O-C})_{\rm RV}
  \left( \frac{M_\star a_b}{G} \right)^{1/2}
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2}
  \nonumber
  \\
  M_{\rm c} &< 
  23.8\,M_\oplus
  \times 
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2},
  \label{eq:rv_bound}
\end{align}
for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto \sin^2 i_{\rm c}$
a geometric prefactor that depends on the argument of periastron
$\omega_{\rm c}$ and inclination $i_{\rm c}$ of the exterior companion
(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b
transits along the line of sight, $f$ can be arbitrarily small, and
both of the preceding limits can be satisfied.

@@

Another mechanism to produce the observed period change would be if an
unseen companion in the WASP-4 system accelerated the star towards our
line of sight.
This would produce a period derivative
\begin{equation}
	\dot{P} \approx \frac{\dot{v}_{\rm r} P}{c},
\end{equation}
for $\dot{v}_{\rm r}$ the time derivative of the radial velocity.
%Recently, \citet{bonomo_gaps_2017} collected all available HARPS data
%for the system, and reported no detectable slope at $3\sigma$.
%However,
\citet{knutson_friends_2014} collected 5 Keck HIRES radial
velocity measurements for WASP-4, and merged them with measurements
from CORALIE and HARPS \citep{wilson_wasp-4b_2008,pont_determining_2011,husnoo_observational_2012}.
Using these datasets, \citet{knutson_friends_2014} found 
\begin{equation}
\dot{v}_{\rm r} =
   -0.0099^{+0.0052}_{-0.0054}
   \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1},
\end{equation}
or that the
star can be accelerating towards us at most by $\dot{v}_{\rm r} > -
0.021 \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$ (2$\sigma$).
The associated $\dot{P}$ must be less than $9\times 10^{-11}$.  
Radial acceleration could thus cause at most one-quarter of the
observed period decrease (Equation~\ref{eq:dP_dt_obs}).
Improved radial velocity
monitoring will help clarify the situation,
particularly given that WASP-4 shows short-term RV trends with
amplitudes $\sim$$\mathcal{O}(10\,{\rm m\,s}^{-1})$, over tens of days
\citep{husnoo_observational_2012}.  Given that WASP-4 is a
magnetically active star \citep{sanchis-ojeda_starspots_2011}, these
short-timescale variations might conspire to produce spurious
RV signals over long baselines.
